COVID19 - Forecast analysis

### 2020-03-13 draft_analysis_Italy.Rmd
Add home page, navigation and theme

```{r  plot data, echo=TRUE}
plot(dat_csv$data,dat_csv$totale_attualmente_positivi,ylab="Total Covid cases",xlab="Date")
mean(diff(log(dat_csv$totale_attualmente_positivi)))
```
The plot shows an exponential gros. We estimate the R0 parameter by means of a linear model. R0 indicates how contagious an infectious disease is. It's also referred to as "the reproduction number" of COVID19.

@@

where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.  
We calculate several R0 values, each one based on a different number of days before the last day. 
The R0 shows a decreasing trend in the last period. The slope b indicates the rate of exponetial increase.  
However we considere the value of the last 10 days, because is more stable.
R code:  
```{r data}
head(dat_csv)
fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(days-10):days,])
summary(fit1)
```

@@

The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.
In the calculation we considered an incubation period of 14 days for two reasons: 1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14. A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to consider in this phase.
```{r r0 plot, echo=TRUE}
slope <-coef(summary(fit1))[2,1]; slope
slope.se <- coef(summary(fit1))[2,2]; slope.se
### R0 estimates and 95% IC 
R_0=slope*14+1;R_0
(slope+c(-1,1)*1.96*slope.se)*14+1
```
We want to make a short term forecast (14 days) with 3 scenario:  
-Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)
-Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% exposed people)
-Scenario 3: 5 exposed people for each COVID-19 case and beta reduced of 50%  (-50% both exposed people and COVID19 contagious power)

>>>>

The actual status is

```{r plots}
library(tidyr)
library(dplyr)
library(ggplot2)
df <- dat_csv %>%
  select(data, ricoverati_con_sintomi, terapia_intensiva, 
         totale_ospedalizzati, isolamento_domiciliare, 
         totale_attualmente_positivi, nuovi_attualmente_positivi, 
         dimessi_guariti, deceduti, totale_casi) %>%
  gather(key = "variable", value = "value", -data)
head(df, 3)

ggplot(df, aes(x = as.Date(data), y = value)) + 
  geom_line(aes(color = variable), size = 1) 
  ```

The plot shows an exponential grow. We estimate the R0 parameter by means of a linear model. R0 indicates how contagious an infectious disease is. It is also referred to as "the reproduction number" of COVID19.

@@

where $`Y_t`$ is the cumulative number of infected at the time t, while b is beta, the slope of the regression line.  
We calculate several R0 values, each one based on a different number of days before the last day. 
The R0 shows a decreasing trend in the last period. The slope b indicates the rate of exponetial increase.  
However we considere the value of the last 10 days, because is more stable.

R code:  
```{r data}
head(dat_csv)
fit1 <- lm(log(totale_attualmente_positivi)~t,data=dat_csv[(days-10):days,])
summary(fit1)
```

@@

The incubation period for the coronavirus is in mean 5.1 days with a range from 2-14 days.
Please see https://www.worldometers.info/coronavirus/coronavirus-incubation-period/.
In the calculation we considered an incubation period of 14 days for two reasons: 1) the majority of cases is asymptomatic, contagiousness is greater than 5, maybe 14. A minority (who made the swab) will have a duration of about 5 days between the start of contagiousness and swab; 2) 14 days is the worst scenario to consider in this phase.
```{r r0 plot, echo=TRUE}
slope <-coef(summary(fit1))[2,1]; slope
slope.se <- coef(summary(fit1))[2,2]; slope.se
### R0 estimates and 95% IC 
R_0=slope*14+1;R_0
(slope+c(-1,1)*1.96*slope.se)*14+1
```
We want to make a short term forecast (14 days) with 3 scenario:

-Scenario 1: 10 exposed people for each COVID-19 case and beta the same (no restrictions made or even no effects)

-Scenario 2: 10 exposed people for each COVID-19 case and beta reduced of 50% (-50% exposed people)

-Scenario 3: 5 exposed people for each COVID-19 case and beta reduced of 50%  (-50% both exposed people and COVID19 contagious power)
